---
title: "Homework 1"
author: "Matthew Præstgaard"
date: "due 09/26/2024"
output:
  rmdformats::downcute:
    self_contained: true
    default_style: "dark"
    downcute_theme: "default"
    code_folding: "show"
  svglite:
    fig.retina: 2
---

### All code and projects available on https://github.com/matthewep528/Intro-to-R

# This document contains the code for homework 1. The first code chunks include loading required packages, creating vectors of numerical and character types, and combining those vectors into a data frame. Then a function is written which takes a numeric vector as its input and outputs the mean of the numbers.

# Housekeeping

```{r,echo=FALSE,message=FALSE,warning=FALSE}
require(knitr)
require(rmdformats)
knitr::opts_chunk$set(error = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
```

# Task 2: Creating Different Object Types
## Creating numeric vector "my_numbers"

```{r}
my_numbers <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
```

## Creating vector of my five favorate colors

```{r}
my_colors <- c("green", "blue", "red", "purple", "orange")
```

## Creating data frame of name, age, and grade

```{r}
name <- c("Guiseppe", "MargaretThatcher", "Theodore")
age <- c(22, 87, 19)
grade <- c("sophomore", "senior", "junior")

student_info <- data.frame(name, age, grade)
```

# (Task 3: Scripting and Documentation)

# Task 4: Basic Functions

```{r}
# creating function calculate_average using argument "vector" as input
calculate_average <- function(vector) {
    mean(vector)
}

# testing function with "my_numbers"
calculate_average(my_numbers)
```

# Task 5: Good Programming Practices
### I would like to think I have good commenting and legibility practices thanks to the lintr package automatically correcting things as I go