---
title: "Homework 2"
author: "Matthew Præstgaard"
date: "due 10/24/2024"
output:
  rmdformats::downcute:
    self_contained: true
    default_style: "dark"
    downcute_theme: "default"
    code_folding: "show"
  svglite:
    fig.retina: 2
---

### All code and projects available on https://github.com/matthewep528/Intro-to-R

# Housekeeping

```{r,echo=FALSE,message=FALSE,warning=FALSE}
require(knitr)
require(rmdformats)
knitr::opts_chunk$set(error = TRUE)
```

```{r}
# Function to check and install missing packages
checkPackages <- function(packages) {
  for (pkg in packages) {
    if (!requireNamespace(pkg, quietly = TRUE)) {
      install.packages(pkg)
    }
  }
}

# making vector of required packages
requiredPackages <- c("tidyverse", "rio", "expss", "plotly", "ggh4x", "devtools")

# running function and applying
checkPackages(requiredPackages)

lapply(requiredPackages, library, character.only = TRUE)

# Check if 'ggthemr' is installed
if (!requireNamespace("ggthemr", quietly = TRUE)) {
  message("Package 'ggthemr' not found. Installing from GitHub...")

  # Install 'ggthemr' from GitHub
  devtools::install_github("Mikata-Project/ggthemr")
}

# Load 'ggthemr'
library(ggthemr)
```

```{r, message=FALSE}
# require(tidyverse)
# require(rio)
# require(expss)
# require(plotly)
# require(ggh4x)
source(here::here("Homework2", "csvImport.r")) # sourcing script for importing sleep data
```

# Part I: Data Import

```{r}
# script was already sourced, but is also included here in a code chunk
# script to import CSV data set
require(tidyverse)
require(rio)
require(here)

diet <- rio::import(here("Homework2", "diet_data-2.csv"))

# inspecting data
dim(diet)
head(diet)
names(diet)
str(diet)

# creating data set with only adults GE 20 years old
dietAdult <- diet %>%
  filter(age >= 20)
```

# Part II
## Create a new age categorical variable using ten-year intervals, starting at aged 20

```{r}
dietAdult <- dietAdult %>%
  dplyr::mutate(age_category = case_when(
    age >= 20 & age < 30 ~ "20-29",
    age >= 30 & age < 40 ~ "30-39",
    age >= 40 & age < 50 ~ "40-49",
    age >= 50 & age < 60 ~ "50-59",
    age >= 60 & age < 70 ~ "60-69",
    age >= 70 & age < 80 ~ "70-79",
    age >= 80 & age < 90 ~ "80-89",
    age >= 90 ~ "90-99"
  ))
```

## Create a new binary variable to indicate good or bad dietary quality: good dietary quality with healthy eating index (HEI) > 80, bad dietary quality with HEI < 80

```{r}
# using if_else statement to create good and bad categories for diet quality
dietAdult <- dietAdult %>%
  mutate(
    dietQuality = if_else(hei > 80, "Good", "Bad")
  )
```

## Create two new physical activity categorical variables:Met Physical Activity Recommendations (MVPA >150 min/week) or Did Not Meet Physical Activity Recommendations (<150 min/week).Sufficient Physical Activity (MVPA >150 min/week), Insufficient Physical Activity (0< MVPA <150 min/week), and No Reported Physical activity (0 min/week)

```{r}
dietAdult <- dietAdult %>%
  mutate(
    metPhysAct = if_else(mvpa > 150, "Yes", "No"),
    sufficientPhysAct = case_when(
      mvpa > 150 ~ "Sufficient",
      mvpa > 0 & mvpa <= 150 ~ "Insufficient",
      mvpa == 0 ~ "NotReported"
    )
  )
```

```{r,message=FALSE,warning=FALSE}
# ggplot histogram function with optional faceting variable
# Usage:gg.hist(data, x, bins(optional), title(optional), data$group(optional))
ggHist <- function(data, x, bins = 20, title = "Insert Title", x_lab = "x", group = NULL) {
  require(ggplot2)
  require(ggh4x)

  gg <- ggplot(data, aes(x = {{ x }})) +
    geom_histogram(
      aes(
        y = after_stat(density)
      ),
      bins = bins,
      alpha = 0.8,
    ) +
    stat_theodensity(
      col = "#ecd467",
      size = 0.9,
      linetype = "dashed"
    ) +
    labs(
      title = title,
      x = x_lab,
      y = "Frequency"
    )

  if (!is.null(group)) {
    gg <- gg + aes(fill = {{ group }}) +
      facet_wrap({{ group }})
  }

  return(gg)
}
```

```{r, warning = FALSE}
# making some fun plots using some of these newly created variables
ggHist(dietAdult, age, group = dietAdult$metPhysAct, title = "Histograms of Age Grouped by Meeting Physical Activity", x_lab = "Age") %>% ggplotly()
```

## Create dummy variables for each race group

```{r}
dietAdult <- dietAdult %>%
  mutate(
    dRace = (race - 1)
  )
```

## Label both variables and values, use Diet_Data_Codebook. Please also check whether the labels are created correctly.

```{r}
# Applying variable labels
dietAdultLab <- apply_labels(dietAdult,
  age = "Age, yrs",
  edu = "Education level",
  work = "Employment status",
  poverty = "Poverty status",
  mvpa = "Physical activity, min/wk",
  sleep_duration = "Sleep duration, hrs/night",
  eat_window = "Food intake window, hrs",
  eat_window_10hr = "10 hr food intake window",
  hei = "Healthy eating index",
  gender = "Gender",
  bmi = "Body Mass Index",
  race = "Race/ethnicity",
  age_category = "Age Category (10-yr intervals)",
  dietQuality = "Dietary Quality",
  metPhysAct = "Met Physical Activity Recommendations",
  sufficientPhysAct = "Sufficient Physical Activity",
  dRace = "Dummy Variable for Race"
)

# Checking if the labels are applied
sapply(dietAdultLab, var_lab)

# Applying Value Labels
dietAdultLab <- apply_labels(dietAdultLab,
  edu = c("< high school" = 1, "high school/GED" = 2, "> high school" = 3),
  work = c("not working" = 0, "part-time" = 1, "full-time" = 2),
  poverty = c("low income" = 0, "middle income" = 1, "high income" = 2),
  gender = c("Men" = 1, "Women" = 2),
  race = c("non-Hispanic White" = 1, "non-Hispanic Black" = 2, "Hispanic" = 3, "Other (includes multiracial)" = 4),
  metPhysAct = c("Yes" = 1, "No" = 0),
  sufficientPhysAct = c("Sufficient" = 1, "Insufficient" = 2, "NotReported" = 0),
  dietQuality = c("Good" = 1, "Bad" = 0)
)

# Checking if the labels are applied again
sapply(dietAdultLab, var_lab)
```